Active contours-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-09-07 Publications/2015-NeuroImage/expanded.tex
add expanded

Early attempts to delineate the \gls*{wm} involved thresholding the 
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses are highly dependent on the threshold set
  \citep{taoka_fractional_2009}.
To overcome the unreliability of \gls*{fa} thresholding,
  \cite{zhukov_level_2003} proposed active contours with edges represented
  by level sets, which evolve on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based level sets based on the eigenvalues of tensors fitted using the
  \gls*{dmri} data.
Many studies have addressed the definition of appropriate features for clustering,
  such as the 5D[A1] representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixed models based on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clustering,
  graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

To address the segmentation problem by registration, \cite{saad_new_2009} 
  used Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and 
  the \emph{b0}.
Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
which uses active contours with edges to search for intensity boundaries in the \emph{b0}
  image.
The active contours are initialized using surfaces extracted from the 
  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
  surface (the \gls*{wm}/\gls*{gm} interface). 
The \emph{b0} image only includes a detectable frontier for the pial surface, so 
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
This tool has become the standard method because of its proven robustness, although the 
  distortions found in \gls*{dmri} are nonlinear.
To overcome this issue, \emph{bbregister} excludes the
  regions that are typically warped by artifacts from the boundary search.
However, because the distortion is not considered, it must be addressed separately.
Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
  images based on their similarity as a unique method for correcting distortion
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.

@@

\makeatletter{}\section{Methods}
\label{sec:methods}
Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
  extracted from the undistorted \gls*{t1} image (the reference space $R$).
We reformulate the segmentation of the distorted \gls*{dmri} images (the moving space $M$)
  as a registration problem where we search for an underlying deformation field $U$ such that 
  the structures in $R$ defined by $\gammaset_R$ align optimally with their corresponding
  structures in $M$:

@@

\paragraph*{Segmentation model}
Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
  we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
  from \gls*{dmri} data.
After evaluating several alternative models, we empirically defined a partition \omegaset{}
  according to the following six regions:
  1) thalamus ($\Omega_{Tha}$);
  2) ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
  3) cerebral \gls*{wm} ($\Omega_{WM}$);
  4) brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
	5) cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
	6) cortical \gls*{gm} ($\Omega_{GM}$).
Using tools in \emph{FreeSurfer} and appropriate selections of labels in the 
  \emph{aparc} segmentation released with the \gls*{hcp} data, we extracted the $\gammaset_R$ set for the
  reference surfaces.
The segmentation model corresponding to this partition is shown in \autoref{fig:model}
  and discussed in greater detail in \suppl{section S4}.

@@

\end{document}
[A1]I understand that many of the abbreviations in this manuscript may be common abbreviations, however, depending on the journal to which you are submitting, it may be necessary to provide definitions in accordance with your target journal author abbreviation guidelines.

>>>>

Early attempts to delineate the \gls*{wm} involved thresholding the
  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
  the \gls*{dmri} data.} map.
However, the mask and subsequent analyses are highly dependent on the threshold set
  \citep{taoka_fractional_2009}.
To overcome the unreliability of \gls*{fa} thresholding,
  \cite{zhukov_level_2003} proposed active contours with edges represented
  by level sets, which evolve on directionally invariant scalar maps.
\cite{rousson_level_2004} successfully segmented the corpus callosum with
  region-based level sets based on the eigenvalues of tensors fitted using the
  \gls*{dmri} data.
Many studies have addressed the definition of appropriate features for clustering,
  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
Other efforts include mixed models based on sets of directionally invariant maps
  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
  hierarchical \citep{lu_segmentation_2008} clustering,
  graph-cuts \citep{han_experimental_2009},
  and volume fraction modeling \citep{kumazawa_improvement_2013}.

To address the segmentation problem by registration, \cite{saad_new_2009}
  used the Pearson's correlation coefficient to obtain a linear alignment of the \gls*{t1} and
  the \emph{b0}.
Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
  which uses active contours with edges to search for intensity boundaries in the \emph{b0}
  image.
The active contours are initialized using surfaces extracted from the
  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
  surface (the \gls*{wm}/\gls*{gm} interface).
The \emph{b0} image only includes a detectable frontier for the pial surface, so
  \emph{bbregister} is limited to aligning the cortical layer in this
  application.
This tool has become the standard method because of its proven robustness, although the
  distortions found in \gls*{dmri} are nonlinear.
To overcome this issue, \emph{bbregister} excludes the
  regions that are typically warped by artifacts from the boundary search.
However, because the distortion is not considered, it must be addressed separately.
Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
  images based on their similarity as a unique method for correcting distortion
  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.

@@

\makeatletter{}\section{Methods}
\label{sec:methods}
Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
  extracted from the undistorted \gls*{t1} image (the reference space $R$).
We reformulate the segmentation of the distorted \gls*{dmri} images (the moving space $M$)
  as a registration problem where we search for an underlying deformation field $U$ such that 
  the structures in $R$ defined by $\gammaset_R$ align optimally with their corresponding
  structures in $M$:

@@

\paragraph*{Segmentation model}
Based on our experience   and previous studies \citep{ennis_orthogonal_2006},
  we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
  from \gls*{dmri} data.
After evaluating several alternative models, we empirically defined a partition \omegaset{}
  according to the following six regions:
  1) thalamus ($\Omega_{Tha}$);
  2) ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
  3) cerebral \gls*{wm} ($\Omega_{WM}$);
  4) brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
  5) cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
  6) cortical \gls*{gm} ($\Omega_{GM}$).
Using tools in \emph{FreeSurfer} and appropriate selections of labels in the 
  \emph{aparc} segmentation released with the \gls*{hcp} data, we extracted the $\gammaset_R$ set for the
  reference surfaces.
The segmentation model corresponding to this partition is shown in \autoref{fig:model}
  and discussed in greater detail in \suppl{section S4}.

@@

\end{document}
